---
title: "Report"
authors:
output: pdf_document
---

# A Describe  

## 1 Time  

## 2 Age  

## 3 Gender  

## 4 Race  


# B Evaluate  

## Harrassment or Efficiency?  

```{r}
# import variables

library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggrepel)
library(plyr)
library(stargazer)

setwd("~/00. Spring/Business Analytics/Group Project")

stop_and_search <- read_csv("Dataset/2016-07-metropolitan-stop-and-search.csv")
crimes <- read_csv("Dataset/2016-07-metropolitan-crimes.csv")
events_by_postcode <- read_csv("Dataset/events-by-postcode.csv")
postcode <- read_csv("Dataset/postcode.csv")

# describe: create "month" for select July data
stop_and_search <- stop_and_search %>% mutate(month=month(ymd_hms(stop_and_search$Date)))
stop_and_search <- stop_and_search %>% filter(month==7)

# evaluation: create "SSO" for searching outcome
stop_and_search <- stop_and_search %>% 
  mutate(SSO = ifelse(grepl('Nothing', stop_and_search$Outcome), 'Nothing', 'Something'),
         Something = ifelse(SSO =='Something', 1, 0))

# Explain: create "hour" and "time_of_day" for regression
stop_and_search <- stop_and_search %>% mutate(hour=hour(ymd_hms(stop_and_search$Date)),
                                              time_of_day=case_when(hour %in% c(5,6,7,8,9)~"morning",
                                                                    hour %in% c(10,11,12,13,14,15,16)~"noon",
                                                                    hour %in% c(17,18,19,20,21,22)~"evening",
                                                                    hour %in% c(23,0,1,2,3,4)~"midnight")) %>% select(-hour)
stop_and_search$time_of_day <- factor(stop_and_search$time_of_day,levels = c("Morning", "Noon", "Evening", "Midnight"))

data_3 <- stop_and_search %>% filter(!is.na(`Age range`), Gender %in% c("Male", "Female"), 
                                     `Officer-defined ethnicity` %in% c('Black', 'White', 'Asian'))

 # Predict
ggplot(events_by_postcode, aes(x=something, y=Total, col=ratio))+geom_point()

```

```{r}  
# RATIOS
count1 <- length(which(stop_and_search$Something == 1)) 
total_count <- length(stop_and_search$Something)
success_ratio <- count1/total_count * 100
success_ratio

stop_and_search$Controlled_drugs <- ifelse(stop_and_search$'Object of search' == "Controlled drugs", 1, 0)
stop_and_search$Criminal_damage <- ifelse(stop_and_search$'Object of search' == "Articles for use in criminal damage", 1, 0)
stop_and_search$Threaten_or_harm <- ifelse(stop_and_search$'Object of search' == "Anything to threaten or harm anyone", 1, 0)
stop_and_search$Firearms <- ifelse(stop_and_search$'Object of search' == "Firearms", 1, 0)

# Success ratios for different objects of search
drugs_count <- length(which(stop_and_search$Controlled_drugs == 1)) 
damage_count <- length(which(stop_and_search$Criminal_damage == 1)) 
threatening_count <- length(which(stop_and_search$Threaten_or_harm == 1)) 
firearms_count <- length(which(stop_and_search$Firearms == 1)) 
success_drugs_count <- sum(stop_and_search$Something == 1 & stop_and_search$Controlled_drugs == 1, na.rm=TRUE)
success_damage_count <- sum(stop_and_search$Something == 1 & stop_and_search$Criminal_damage == 1, na.rm=TRUE)
success_threatening_count <- sum(stop_and_search$Something == 1 & stop_and_search$Threaten_or_harm == 1, na.rm=TRUE)
success_firearms_count <- sum(stop_and_search$Something == 1 & stop_and_search$Firearms == 1, na.rm=TRUE)
ratio_drugs <- success_drugs_count / drugs_count * 100
ratio_damage <- success_damage_count / damage_count * 100
ratio_threatening <- success_threatening_count / threatening_count * 100
ratio_firearms <- success_firearms_count / firearms_count * 100

mydata <- data.frame(total_count, drugs_count, damage_count, threatening_count, firearms_count)
ratios <- data.frame(success_ratio, ratio_drugs, ratio_damage, ratio_threatening, ratio_firearms)
```


# C Explain  

## 6 Racial Fairness  

```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# import variables
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(stargazer)

stop_and_search <- read_csv("Dataset/2016-07-metropolitan-stop-and-search.csv")

# create variables
stop_and_search <- stop_and_search %>% mutate(month=month(ymd_hms(stop_and_search$Date)))
stop_and_search <- stop_and_search %>% filter(month==7)
stop_and_search <- stop_and_search %>% mutate(hour=hour(ymd_hms(stop_and_search$Date)),
                                              time_of_day=case_when(hour %in% c(5,6,7,8,9)~"Morning",
                                                                    hour %in% c(10,11,12,13,14,15,16)~"Noon",
                                                                    hour %in% c(17,18,19,20,21,22)~"Evening",
                                                                    hour %in% c(23,0,1,2,3,4)~"Midnight")) %>% select(-hour)
stop_and_search$time_of_day <- factor(stop_and_search$time_of_day,levels = c("Morning", "Noon", "Evening", "Midnight"))
stop_and_search <- stop_and_search %>% 
  mutate(SSO = ifelse(grepl('Nothing', stop_and_search$Outcome), 'Nothing', 'Something'),
         Something = ifelse(SSO =='Something', 1, 0))
data_3 <- stop_and_search %>% filter(!is.na(`Age range`), Gender %in% c("Male", "Female"), 
                                     `Officer-defined ethnicity` %in% c('Black', 'White', 'Asian'))
model <- glm(Something~time_of_day+`Age range`+Gender+`Officer-defined ethnicity`, family = 'binomial', data=data_3)
```

```{r result, echo=FALSE, results='asis'}
stargazer(model, header = FALSE,  type = 'latex', title="Regression Results") 
```

Comments:  
In this regression model, we can't find a strong relationship between the outcome and racial background. Holding other factors fixed and use Asian as base group, we found that if the race of people who get stooped is black or white, the odd ratio are only about 0.03 and 0.04. That means if people are black or white, the probability of 'something' only increases by 3% and 4%, compared to Asian. What's more, the p-values are 0.6 and 0.5, which are really insignificant. Therefore, we believe there is no obvious relationship between outcome and race.  

# D Predict  

## 7 Does PredPol really work?  

```{r}
library(readr)
library(dplyr)
library(ggplot2)
postcode <- read_csv("Dataset/postcode.csv")
stop_and_search <- read_csv("Dataset/2016-07-metropolitan-stop-and-search.csv")

postcode2 <- postcode %>% group_by(postcode) %>% summarise(nothing=sum(cntnothing),
                                              something=sum(cntsomething)) %>%
  filter(!is.na(nothing)) %>% mutate(total=nothing+something,
                                     ratio=something/total)

# write_csv(postcode2, 'postcode_final.csv')
postcode3 <- postcode2 %>% filter(ratio!=0, ratio!=1, total<600, total>50)

ggplot(postcode3, aes(x=something, y=total, col=ratio)) + geom_point(size=2) +
  scale_colour_gradient(low = "yellow", high = "red")

ggplot(postcode2, aes(x=log(total), y=ratio)) + geom_point(size=2) + geom_smooth()
```

```{r}
# Question d)
postcode <- ddply(postcode,"postcode",numcolwise(sum))
postcode$total <- postcode$cntnothing + postcode$cntsomething
postcode$ratio <- postcode$cntsomething / postcode$total
postcode <- postcode[order(postcode$ratio),]
postcode <- postcode[, -c(2:5)]
postcode1 <- postcode %>% filter(ratio!=0, ratio!=1, total<600, total>50)

ggplot(postcode1, aes(x=cntsomething, y=total, col=ratio)) + geom_point(size=2) +
  scale_colour_gradient(low = "yellow", high = "red") +
  xlab("something")

ggplot(postcode, aes(x=log(total), y=ratio)) + geom_point(size=2) + geom_smooth()
```



# E Automate  

## 8 Startup Pitch  

